---
title: "plsr_neon_plot_traits"
author: "Sarah Graves"
date: "June 23, 2016"
output: html_document
---

# Goal  

Use PLS regression to relate foliar nitrogen (plot level kg/ha) to remote sensing data.

## PLSR model

Decided to use the `PLS` package because it was cited in this paper: http://onlinelibrary.wiley.com/doi/10.1890/13-2110.1/full

```{r SET-UP}

library(pls)
library(ggplot2)
library(plotrix)
library(dplyr)
library(reshape2)

```


# Use Kunxuan's code to create spectra for each plot

```{r, LOAD-HS,eval=FALSE}

# need Leah's code
library(raster)
library(rhdf5)
library(neonAOP)

f <- "../NEONdata/D03-Florida/OSBS/2014/spectrometer/reflectance/Subset1NIS1_20140507_150703_atmcor.h5"

bands <- c(1:426)

# create stack from all bands
# this takes a few minutes to run
hs <- create_stack(f,bands=bands,epsg=epsg)

# pull out pixels to use
# this is just the first 100 pixels (the first 100 cols in the first row)
hs_refl <- hs[1:100,1,1]

# calculate mean spectra for all pixels in raster stack
hs_mean <- extract_av_refl(hs,aFun=mean)

# remove bad bands
aop_bands <- read.csv("supplemental_data/neon_aop_bands.csv")

hs_mean[aop_bands$noise == 1] <- NA

plot(hs_mean)

```

```{r LONG-FUNCTION}

# function to take refl matrix and transform to long format
# set columns as numbers

refl_melt <- function(refl_data.frame,id.var.name){
  
  # assume wavelenghts as column names
  wavelength <- as.numeric(substr(x=colnames(refl_data.frame),
                                  start=4,
                                  stop = nchar(colnames(refl_data.frame))))
  # reassign column names
  colnames(refl_data.frame) <- wavelength
  
  df.long <- melt(refl_data.frame,id.vars=id.var.name)

  # variables are numbers, convert to numeric for scatterplot
  df.long$variable <- as.numeric(as.character(df.long$variable))
  df.long$value <- as.numeric(df.long$value)

}
```



```{r PLOT-SPEC}

hs_refl <- read.csv("output-data/plot_mean_reflectance.csv")

# set columns as numbers
wavelength <- as.numeric(substr(colnames(hs_refl[-1]),4,nchar(colnames(hs_refl)[-1])))  
colnames(hs_refl)[-1] <- wavelength


df.long <- melt(hs_refl,id.vars="plotID")

# variables are numbers, convert to numeric for scatterplot
df.long$variable <- as.numeric(as.character(df.long$variable))
df.long$value <- as.numeric(df.long$value)

ggplot(df.long,aes(x=variable,y=value))+
  geom_point() +
  theme_bw() +
  scale_y_continuous(limits=c(0,.5)) +
  labs(list(title="Plot mean reflectance\n Ordway-Swisher Bio. Station, FL",x="Wavelength",y="reflectance"))

```


## N content data

I will use plot-level kg of N. I am waiting for Jeff to get this? I can also work on it.
For now, I will create fake N data. Because I only have 1 plot, I will create fake N data for all pixels in the h5 clip.

```{r N-DATA,eval=FALSE}

# how many "plots" do I have?
# use the first row of the hs data
dim(hs_refl)

# fake N data
n_kg <- round(runif(n=dim(hs_refl)[1],
              min=5,
              max=100),2)

n_kg

```

```{r N-DATA-REAL}

n_kg <- read.csv("supplemental_data/osbs_plot_n.csv")
n_kg <- n_kg[,-1]

```


## Merge data
```{r MERGE-DATA}

# merge refl data to n data by plot
# keep only those plots with N data
df <- merge(n_kg,
            hs_refl,
            by.y="plotID",
            by.x="plot_id",
            all.x=T,
            all.y=F)

df.use <- data.frame(N=df$foliar_n_per_tree,
                 refl=I(as.matrix(df[,6:374])))

```


```{r TRAIN-TEST-GROUPS}

# don't have enough data to do test or train groups

```

```{r PLS-MODEL}

# build PLSR model
# number of components is limited because there are only 7 plots
m1 <- plsr(N ~ refl, data=df.use, ncomp = 5, validation = "LOO")


summary(m1)


plot(RMSEP(m1), legendpos = "topright")

# plot measured vs. predicted
plot(m1, ncomp = 5, asp = 1, line = TRUE)

plot(m1, "loadings", comps = 1:2, legendpos = "topright", xlab = "nm")
abline(h = 0)

plot(m1, ncomp = 2, asp = 1, line = TRUE,
     main="Plot-level foliar N content (kg)",
     pch=21,
     bg="springgreen",
     cex=2)


```

## Apply model to training data

```{r APPLY-MOD, eval=F}

m1_predict <- predict(m1, ncomp = 3, newdata = dTest)

plot(m1_predict ~ dTest$N,
     main="Predicted vs. Observed")
abline(0,1)

```

